---
title: "Simple Linear Regression"
author: "Yelsh Gebreselassie"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
date: "2022-09-21"
---




In this document, we do basic descriptive and regression analysis to 
understand what variables best predict freshman GPA. 


1. First install/ load the R packages we need. We also load and attach our data. 


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(modelsummary)

gpa.data <- read_csv("data/satgpa.csv")
attach(gpa.data)
```

# Exploratory questions

## How well do SAT scores correlate with freshman GPA?

```{r}
cor(gpa_fy, sat_total) 
```
A correlation of -1 means perfect negative correlation. A correlation of 0 means, no correlation between the two. And a correlation of 1 means perfect positive correlation. The above result shows a positive correlation between SAT scores and freshman GPA.But it is not very strong, meaning close to 1. 

```{r}
ggplot(data = gpa.data, mapping = aes(x = sat_total, y = gpa_fy)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(y = "Freshman GPA", x = "SAT total")
```

The above plot shows the positive correlation between SAT scores and
freshman GPA. 

## How well do high school GPA correlate with freshman GPA?

```{r}
cor(gpa_fy, gpa_hs) 
```

```{r}
ggplot(data = gpa.data, mapping = aes(y = gpa_fy, x = gpa_hs)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "High school GPA", y = "Freshman GPA")
```

## Is the correlation between SAT scores and freshman GPA stronger for men or for women?

```{r}
gpa.data %>% 
  group_by(sex) %>% 
  summarize(correlation = cor(sat_total, gpa_fy))
```

```{r}
ggplot(data = gpa.data, mapping = aes(y = gpa_fy, x = gpa_hs, color = sex)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "High school GPA", y = "Freshman GPA")
```

# Models

## Do SAT scores predict freshman GPAs?

- X = SAT scores
- Y = Freshman GPA

```{r}
model_simple <- lm(gpa_fy ~ sat_total, data = gpa.data)
tidy(model_simple, conf.int = TRUE)
glance(model_simple)
```

## Does a certain type of SAT score have a larger effect on freshman GPAs?

```{r}
model_sat <- lm(gpa_fy ~ sat_verbal + sat_math, data = gpa.data)
tidy(model_sat, conf.int = TRUE)
glance(model_sat)
```

## Do high school GPAs predict freshman GPAs?

```{r}
model_hs <- lm(gpa_fy ~ gpa_hs, data = gpa.data)
tidy(model_hs, conf.int = TRUE)
glance(model_hs)
```

## Explaining College GPA using SAT scores and sex

```{r}
model_sat_sex <- lm(gpa_fy ~ sat_total + sex, data = gpa.data)
tidy(model_sat_sex, conf.int = TRUE)
glance(model_sat_sex)
```

## Explain college GPA using SAT scores, high school GPA and sex

```{r}
model_final <- lm(gpa_fy ~ sat_verbal + sat_math + gpa_hs + sex, 
                  data = gpa.data)
tidy(model_final, conf.int = TRUE)
glance(model_final)
```

## Which model best predicts freshman GPA?

```{r}
modelsummary(list(model_simple, model_sat, model_hs, 
                  model_sat_sex, model_final))
```

