---
title: "Regression"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
date: "2022-09-21"
---

In this document, we do basic descriptive and regression analysis to 
understand what variables best predict freshman GPA. 


1. First install/ load the R packages we need


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(modelsummary)

gpa.data <- read_csv("data/satgpa.csv")
attach(gpa.data)
```

# Exploratory questions

## How well do SAT scores correlate with freshman GPA?

```{r}
cor(gpa_fy, sat_total) 
```
A correlation of -1 means perfect negative correlation. A correlation of 0 means, no correlation between the two. And a correlation of 1 means perfect positive correlation. The above result shows a positive correlation between SAT scores and freshman GPA.But it is not very strong, meaning close to 1. 

```{r}
ggplot(data = gpa.data, mapping = aes(x = sat_total, y = gpa_fy)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(y = "Freshman GPA", x = "SAT total")
```

The above plot shows the positive correlation between SAT scores and
freshman GPA. 
